
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Introduction

This course is meant to get you started writing scientific manuscripts using LaTeX in Overleaf.

## Motivation

Overleaf is a very useful tool for collaborative scientific writing. It has an advantage over other potential collaborative writing options in that it enables for seamless changes in formatting and arrangement of text to allow you to submit to various publication platforms, such as preprint archives or scientific journals. Another advantage of Overleaf is that it is based on using LaTeX which makes it easy to add mathematical formulas to your writing. It also helps you do really complicated formatting and layouts with much more ease than other options and it makes it easy to replicate that complicated formatting/layout again for other papers or other writing products. It is widely used in certain fields like statistics and computer science. Therefore, it can also be useful to know about if you have collaborators in these fields.

## Target Audience  

The course is intended for scientific writers who might be interested in learning how to automate some of their writing formatting/layouts.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "for indiviudals who: Are unfamiliar with what LaTeX is, but like the idea of more customizable and reproducible formatting  of text in documents. Have heard of LaTeX but don’t know how to use it in practice. Want guidance on collaborating with Overleaf."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1bb9ca840c8_0_164")
```


## Curriculum  

The course covers basic information about why LaTeX can be useful, how to get started in overleaf using LaTeX with a template, how to work with a team on Overleaf, and what to do when you encounter problems. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Concepts covered in this course are: 1. What typsetting is, 2. What LaTeX is and how it came to be, 3. How Overleaf makes writing LaTeX documents easier, 4. How to work with a template for a scientific article intended for arXiv, 5. How to collaborate on Overleaf, 6. How to get help if you run into issues"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1bb9ca840c8_0_179")
```



## Objectives

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "This course will demonstrate how to: 1.Be familiar with what LaTeX is and understand the origins of how it came to be, 2. Understand the benefits of using LaTeX and Overleaf, 3. Get started writing in Overleaf with a template for a scientific manuscript, 4. Add references to a document in Overleaf, 5. Get help if you encounter challenges, 6. Work collaboratively on Overleaf"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1bb9ca840c8_0_196")
```



