
# Writing with Overleaf

Now that you know the advantages of using overleaf and have started working with a template, we will now discuss more about how to make additional modifications to your document.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning Objectives: 1.Recognize that there are packages for LaTex, 2. Understand that a command can either change the text within the brackets or cause a change to the overall document or section, 3. Be familiar with some common commands and notations for using LaTex to perform tasks such as:  creating headings, creating comments, starting and ending elements within the text"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_125")
```

## LaTeX Basics

Since we are working with a template, it isn't necessary to learn everything there is to know about writing in LaTeX to get started. However, it can be useful to understand some fundamentals, so we will go over most of the commands that you see listed in the template.

### Document Class 

At the top of the template you will notice `\documentclass{article}`.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Documentclass command"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_7")
```


This specifies general typesetting information about the type of document that we intend to make. For example, it often specifies font size, the overall layout of the text, and alignment of various features of the text. Since we are writing a scientific article, the specification here is `article`.

To learn more about document classes see this [documentation link](https://libguides.utsa.edu/c.php?g=522165&p=3570198) form @wu_libguides.


### Packages

Next you will see that  `\usepackage{}` is repeated several times with different information in the brackets:

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The top of the template shows usepackage several times."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c84a339a79_262_0")
```

We will refer to these tags with brackets such as `\usepackage{}` as **commands** from now on (as this is what the are generally referred to) and they cause a change to either the text within the brackets or the overall document.

The `\usepackage{}` command installs packages, which are bits of  code that help you do additional things with your documents. Packages need to be installed before commands from those packages can be called. Commands from these packages will be utilized later in the template. **It is recommended that you leave this code as is, and only modify the rest of the template until you learn more.** 

In addition to determining what commands you can use, packages will also determine how the content is formatted or laid out.


### Author section

You may recall that we previously described how to bold font using `\textbf{bold text}`.
With LaTeX you will be using brackets often to designate what to do with a specific set of text that is contained within the brackets.

<div class = "warning">
If you do not close a set of brackets you will get an error, so be careful about this.
</div>

As we scroll down the template, we next see the `\title` command that we previously worked with when we modified the text within the brackets to change the title. 


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Recap of the title command"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_12")
```


Then we see the `\author` command, which adds authors to the paper. These will be formatted in the way that is shown on the template. When you see `\\` two backslashes, this indicates that the line is finished and a new one is to be made. For tables or formatting like the authors, it should work well, but it is best to  **[avoid](https://tex.stackexchange.com/questions/225893/what-does-double-backslash-in-latex-mean)** using this for line breaks within the paragraphs that you might include in the paper. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Double backslash ends a line."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_45")
```


We also see another command `\textttt{}` used within the `\author{}` command to change the text to typewriter font.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The texttt command changes the font within the brackets to typewriter font."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_67")
```

You may also notice `%% examples of other authors` is in green and does not show up in the rendered document. This is what is called a **comment** and it can be used to write notes about the material.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The double percent sign %% creates a comment that will not show up in the final document."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_79")
```

In the author section, the `\And` allows for additional authors to be added. It must be used between each author listing. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "In the author section, the And tag allows for additional authors to be added and needs to be used between each other."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_31")
```

Finally, the author section needs to be completed by closing the brackets.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "In the author section, the brackets for the original author command need to be closed."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_93")
```

### Formatting the Document Body

The command `\begin{document}` enables us to format text for the body of the article. This command comes directly after the author section. It will be paired with `\end{document}` that you will see at the bottom of the template if you scroll down. If you put `%%` in front of the command to change it to a comment (therefore not using it), you will see that the overall document is formatted slightly differently. 


The `\maketitle` will add the title the page where the `begin{document}` command was used. If it is placed before `begin{document}`, the title and authors will show up on a separate page. You can test moving this command around to see how the document changes.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The begin and maketitle commands specify where certian formatting should occur."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_101")
```

### Abstract

The abstract section can be distinguished using the `begin{}` and `end{}` functions just like we used for the body of the document. These two commands will also be used later to indicate that a specific part of the document has started or ended.

### Dummy text

You may notice `\lipsum[]` is used to create random chunks of text. The number within the brackets indicates what specific dummy paragraph to use; the maximum value allowed in the brackets is 150.

### Sections

- `\section{section name}` - This will help you to create sections in the template. We don't need to do anything to modify the text, it will automatically bold the text and number the sections (1, 2, 3 etc.).

- `\subsection{subsubsection name}` - This will help you to create subsections. These headings will be one level down from the section headings and will be numbered like 1.1, 1.2. 

- `\subsubsection{subsubsection name}` - This will help you to create sections one level down from subsections. These heading will be numbered like 1.1.1, 1.1.2.


In the template you can see how these are formatted:

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of sections in the template showing how it creates a number and bold text"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_0")
```

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of a subsection with dummy text."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_110")
```


You can specify if you want sections to be not be numbered when using an asterisks `*` between the command name and the brackets `section*{}`. 

The `\paragraph{}` command works similarly, but drops it will not be numbered. The text within the brackets is an optional word that will be bolded to start the paragraph. You can also leave it empty. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "A paragraph section will not be numbered but whatever text is within the brackets will be in bold."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_118")
```



### Equations

It can be very helpful to include a mathematical equation. To do so we need to use our handy `\begin{equation}` and `\end{equation}` functions to indicate the boundaries. Using `equation` within the brackets indicates that this should be formatted in a certain way. It will center the text nicely and number it.

For more information about mathematical expressions in overleaf see the [Overleaf documentation](https://www.overleaf.com/learn/latex/Mathematical_expressions).

## Conclusion

We hope that this chapter has helped you feel more confident to start filling out the template with your own content.

Here are some of the major take-home messages:

1. The `\documentclass{article}` command indicates that we are creating a scientific article.
1. Packages are used to add additional features for users, to install them for a document we need to use the `\usepackage{}` command.
1. Be careful about your brackets when writing your code. If you leave one set open it could cause an error.
1. The double backslash `\\` can indicate the end of an item in the author list for example.
1. The `begin{}` and `end{}` commands indicate when an element is starting or finishing. This could just be an abstract, or the entire body of the document.
1. Section headers can be automatically numbered and styled. The `section{}, `subsection{}`, and `subsubsection{}` will make progressively more nested sections.
1. LaTeX is great for including mathematical notations and equations in documents.

