
# Writing with Overleaf

Now that you know the advantages of using overleaf and have started working with a template, we will now discuss more about how to make additional modifications to your document.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning Objectives: 1. Recognize some basic LaTeX tags, 2. Recognize how Overleaf offers extra support, 3. Add images to documents, 4. Add references to documents, 5. Identify when there is an Issue. 6. Know how to get help"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c84a339a79_0_39")
```

## LaTeX Basics

Since we are working with a template, it isn't necessary to learn everything there is to know about writing in LaTeX to get started. However, it can be useful to understand some fundamentals, so we will go over everything that you see listed in the template.

### Document Class 

At the top of the template you will notice `\ducumentclass{article}`.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Documentclass command"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_7")
```


This specifies general typesetting information about the type of document that we intend to make. For example, it often specifies font size, the overall layout of the text, and alignment of various features of the text. Since we are writing a scientific article, the specification here is `article`.

To learn more about document classes see this [documentation link](https://libguides.utsa.edu/c.php?g=522165&p=3570198) form @wu_libguides.


### Packages

Next you will see that  `\usepackage{}` is repeated several times with different information in the brackets:

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The top of the template shows usepackage several times."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c84a339a79_262_0")
```

We will refer to these tags with brackets such as `\usepackage{}` as **commands** from now on (as this is what the are generally referred to) and they cause a change to either the text within the brackets or the overall document.

The next commands install some packages. Packages are additional code to help you do additional things with your documents. If a command comes from a specific package, you will need to install the package that it comes from first. Commands from these packages will be utilized later in the template. It is recommended that you leave this code as is, and only modify the rest of the template until you learn more. 

In addition to determining what commands you can use, packages will also determine how the content is formatted or laid out.


### Layout Commands

You may recall that we previously described how to bold font using `\textbf{bold text}`.
With LaTeX you will be using brackets often to designate what to do with a specific set of text that is contained within the brackets.

Next, we see the `\title` command that we previously worked with, as we already modified the text within the brackets to change the title. 


Next we see the `\author` command to add authors to the paper formatted in the way that is shown on the template. When you see 

Other important commands will be:

- `\section{section name}` - This will help you to create sections in our template. We don't need to do anything to modify the text, it will automatically bold the text and number the sections (1, 2, 3 etc.).

- `\subsection{subsubsection name}` - This will help you to create subsections, these headings will be one level down from the section headings and will be numbered like 1.1, 1.2. 

- `\subsubsection{subsubsection name}` - This will help you to create sections one level down from subsections, these heading will be numbered like 1.1.1, 1.1.2.


In the template you can see how these are formatted:

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of sections in the template."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_0")
```

The paragraph