
# Figure, tables, and citations oh My!

No that you know some of the basics about how to add text to a template, we want to help you add other elements that are important for scientific communication, like adding figures and images, adding tables, and adding references/citations. In this next chapter, we will introduce how to do this in LaTeX/Overleaf.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning Objectives: 1. Add figures and images to documents, 2. Add references and citations to documents, 3. Add tables to documents, 4. Create reference links back to a figure, table, or section"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_131")
```

You will notice that the template has a few examples of each of these elements and we will walk through these now. 

## Figures and Images

Most scientific articles have figures, so it is helpful to know how to add these to documents. There are two main steps involved in this once you have a figure file that you are ready to add, say a png or a jpeg file.

When you add a figure in LaTex you will use the `\begin{}` and `end{}` functions just like you have done previously. However, this time the you will use `\begin{figure}` and `\end{figure}` to tell Overleaf that this is the type of element you are creating.

If you focus on the second figure in the template, which is probably all you will need to learn (the first figure is a box created in LaTeX), then you will notice that in addition to the `\begin{figure}` and `\end{figure}`  commands, there is a `\centering{}`command, which will align the figure to the center of the page. The `includegraphics{text.png}` portion of the code actually adds the figure into the document. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of the code for a figure when you have a png file to use."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_4")
```

<div class = "notice" > 
Note that comments with one percent sign can be used after code - thus the `% picture` is just telling you that this part of the code is adding a picture. People use different numbers of percent signs based on preference and convention, but just one percent sign is sufficient to turn anything following that into a comment. 
</div>
 
You may notice on the left side of the template in overleaf that there are a few files listed, including the name of the image file used in the code for the figure: `test.png`. The template.tex file is the file we have been working in.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The template.tex file is the file we have been working in. You can see it in the left menu."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_9")
```

If you click on the name of the image file you will see a preview of the image.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Location of the image file called test.png in the left menu."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_19")
```

<details><summary> **What are these other files?**</summary>
<div class = "notice">
The **template.tex** file is the file we have been working in that contains the main document  content text and code to format the text.
The **`reference.bib`** file contains the bibliography information that will be used by the template.tex file.
The **`README.md`** file will tell you more about the template that you are using. <br>
The **`arxiv.sty`** file is a style file that contains code more specifically style the document  for arXiv preprints. This code is then applied in our `template.tex` file (the one we have been working in) by the command \usepackage{arxiv}, because that is the name of the `.sty` file. 
</div>
</details>


To upload a new image to add a new figure, you can click on the new file button which has an icon that looks like a piece of paper with the right upper corner folded.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "New file button"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_27")
```

Then select the upload button to drag and drop a new image file from your computer.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Menu for uploading a file.."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_33")
```

Then to add this new figure to your document you would just need to modify the code above so that the name of the image file matched you uploaded.

`\begin{figure} % picture`
   ` \centering`
   ` \includegraphics{new.png}`
`\end{figure}`


## Creating interal links

You may have noticed a command `\label{}` when looking through the template.
This is a very helpful command that helps you to create a tag to refer back to when you have a figure or a section header that you want to refer to. There is a nifty command `\ref{}` that helps you create these references. However, it is a tiny bit tricky, so we will walk through a couple of examples.

