
# Figure, tables, and citations oh My!

No that you know some of the basics about how to add text to a template, we want to help you add other elements that are important for scientific communication, like adding figures and images, adding tables, and adding references/citations. In this next chapter, we will introduce how to do this in LaTeX/Overleaf.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning Objectives: 1. Add figures and images to documents, 2. Add references and citations to documents, 3. Add tables to documents, 4. Create reference links back to a figure, table, or section"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_131")
```

You will notice that the template has a few examples of each of these elements and we will walk through these now. 

## Figures and Images

When you add a figure in LaTex you will use the `\begin{}` and `end{}` functions just like you have done previously. However, this time the you will use `\begin{figure}` and `end{figure}` to tell Overleaf that this is the type of element you are creating.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of the code for a figure when you have a png file to use."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_4")
```
 


Once you have the code for actually creating the figure, you then need to put it somewhere in the text. 


You may have noticed a command `\label{}` when looking through the template.
This is a very helpful command that helps you to create a tag to refer back to when you have a figure or a section header that you want to refer to. There is a nifty command `\ref{}` that helps you create these references. However, it is a tiny bit tricky, so we will walk through a couple of examples.

