
# Figure, tables, and citations oh My!

No that you know some of the basics about how to add text to a template, we want to help you add other elements that are important for scientific communication, like adding figures and images, adding tables, and adding references/citations. In this next chapter, we will introduce how to do this in LaTeX/Overleaf.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning Objectives: 1. Add figures and images to documents, 2. Add references and citations to documents, 3. Add tables to documents, 4. Create reference links back to a figure, table, or section"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1d07d88c199_0_131")
```

You will notice that the template has a few examples of each of these elements and we will walk through these now. 

## Figures and Images

When you add a figure in LaTex you will use the `\begin{}` and `end{}` functions just like you have done previously. However, this time the you will use `\begin{figure}` and `\end{figure}` to tell Overleaf that this is the type of element you are creating.

If you focus on the second figure in the template, which is probably all you will need to learn (the first figure is a box created in LaTeX), then you will notice that in addition to the `\begin{figure}` and `\end{figure}`  commands, there is a `\centering{}`command, which will align the figure to the center of the page. The `includegraphics{text.png}` portion of the code actually adds the figure into the document. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example of the code for a figure when you have a png file to use."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_4")
```
 
You may notice on the left side of the template in overleaf that there are a few files listed.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The template.tex file is the file we have been working in. You can see it in the left menu."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c91f28435b_0_9")
```




Once you have the code for actually creating the figure, you then need to put it somewhere in the text. 


You may have noticed a command `\label{}` when looking through the template.
This is a very helpful command that helps you to create a tag to refer back to when you have a figure or a section header that you want to refer to. There is a nifty command `\ref{}` that helps you create these references. However, it is a tiny bit tricky, so we will walk through a couple of examples.

