
# Troubleshooting

Overleaf has lots of ways to make writing in LaTeX easier. In this chapter we will cover several of these ways, as well as how to identify and fix a problem if you encounter one.

## Code Suggestions

As you type code, Overleaf makes code suggestions based on what you have already typed. You can click on a suggestion to use it. Overleaf will also suggestion what to put within command brackets. This can be great if you can't quite remember what a command is or if you are trying to write a document faster.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "TOverleaf makes code suggestions based on what you have already typed. You can click the suggestion to autocomplete your code."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c917691280_100_0")
```

## Symbol Palette

If you need to use many symbols or mathematical notations in your work, the symbol palette can be very helpful. You can access it by pressing the button that looks like an omega symbol at the top of the source editor.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The symbol palette can be very helpful. It can be accessed at the top of the source editor. "}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c917691280_100_7")
```

https://tex.stackexchange.com/

https://www.overleaf.com/learn/how-to/LaTeX_checklist_for_arXiv_submissions

https://arxiv.org/help/submit_tex
https://nasa.github.io/nasa-latex-docs/html/examples/table.html
