
# Troubleshooting

Overleaf has lots of ways to make writing in LaTeX easier. In this chapter we will cover several of these ways, as well as how to identify and fix a problem if you encounter one.

## Code Suggestions

As you type code, Overleaf makes code suggestions based on what you have already typed. You can click on a suggestion to use it. Overleaf will also suggestion what to put within command brackets. This can be great if you can't quite remember what a command is or if you are trying to write a document faster.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "TOverleaf makes code suggestions based on what you have already typed. You can click the suggestion to autocomplete your code."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c917691280_100_0")
```

## Symbol Palette

If you need to use many symbols or mathematical notations in your work, the symbol palette can be very helpful. You can access it by pressing the button that looks like an omega symbol at the top of the source editor.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The symbol palette can be very helpful. It can be accessed at the top of the source editor. "}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1c917691280_100_7")
```


## History

If you change something and want to see what it was like previously, you can checkout the History tab. 

Here you will see the most recent changes as slices of changes overtime. Changes will be highlighted in a color that coordinates with specific users. So if you have collaborators you can see their changes here as well.

If you are using the comparison mode you will see the changes and who made them. However, you can also just look at the files from a specific point in time without seeing changes by clicking on the "view single version" button.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The History Tab shows you changes when in comparison mode. Changes are highlighted in color."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1cc5927a0b3_0_12")
```

For more information about using the history features of Overleaf, see this [documentation](https://www.overleaf.com/learn/latex/Using_the_History_feature) @overleaf_history.

## Log

If you check the log by clicking on the button to the right of the recompile button, you will see messages about any warnings or errors.

The link button at the top of each message will take you to the line of the file that is problematic. 

Hovering on the red circle right next to the line number will show you information about what might be wrong with that line.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Log errors can take you to the line that is problematic"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1UgGtVn7RsqdQ4pJxDk_dueSyREHcH-uWTNAT27E2mG8/edit#slide=id.g1cc5927a0b3_0_17")
```




https://tex.stackexchange.com/

https://www.overleaf.com/learn/how-to/LaTeX_checklist_for_arXiv_submissions

https://arxiv.org/help/submit_tex
https://nasa.github.io/nasa-latex-docs/html/examples/table.html
